### 2024-02-20 appendices/app_taxonomy.tex
First part of intro: transition, VRES, electrofuels, ESOMs

First and foremost, in the context of the energy transition, fuels will keep playing a major role in the global energy system \cite{Ahlgren2012}. Even if electricity gains shares through the electrification of the energy demand, it will not entirely displace fuels for three main reasons: storage, infrastructure compatibility and cross-sectorial links. Because of their intermittency and space disparity, a deeper integration of variable renewable energy sources (VRES) requires storage and transport in order to supply the energy demand at the right time and in the right place \cite{Brouwer2016,Evans2012,Gallo2016,hall2008,Rosa2017}. Where typical container of batteries are limited in terms of storage capacity (up to 10~MWh) and present significant costs and self-discharge losses, the energy conversion into fuels provides a more affordable solution for higher storage capacity (from 100~GWh) and longer storage time scales (months to year) \cite{Rosa2017}. Due to the economic inertia and their infrastructure legacy \cite{Ahlgren2012}, fuels remain the most appropriate solution for sectors requiring high energy density (\eg heavy-duty transport, shipping, aviation or chemical industry) \cite{Albrecht2020,Decker2019,Goede2018,Pearson2012,Rosa2017,Stancin2020,Trieb2018,Zeman2008}. 
\citet{contino2020whole} point out that the energy transition is an interdisciplinary effort and not solely about the power sector. The latter represents only a fifth of the global energy consumption \cite{iea_renewable}. Also, \citet{Goede2018} has shown in 2018 that \ce{CO2} emissions of the Netherlands are equally apportioned between the different types of end-use-demands (\ie power, heat, mobility and non-energy). This highlights the necessity to consider every energy sector rather than focusing all efforts on the power system and even more, to shift towards a multi-vector interconnected energy system. In this cross-sectorial approach and with the perspective of increasing shares of VRES, fuels are promising energy carriers in order to maximise the overall system efficiency \cite{mathiesen2015, Stancin2020}.

@@

Where terms like ``synthetic gas", ``electrofuels" or ``synthetic natural gas" account for 19\%, 13\% and 8\% of the additional relevant literature, respectively, ``synthetic fuels" is used in 35\% of the case to cover fuels produced from a wide variety of primary resources (renewable or not) and through various conversion processes. To pick only a few examples, its usage ranged from liquid hydrocarbons produced from synthesis gas through the Fischer-Tropsch process \cite{haarlemmer2014,Trieb2018} to ammonia from the Haber-Bosch process \cite{bargiacchi2019} or biomass gasification or pyrolysis \cite{monaco2018, rao2015}. As a consequence of an increasing share of VRES in the energy system, ``electrofuels" are gaining more popularity: where this term was used in 4\% of the relevant literature between 2006 and 2014 \cite{ridjan2016}, it is now present in 13\% of the additional relevant literature reviewed for this work. These fuels are considered as being produced mainly based on the electricity provided by VRES \cite{brynolf2018}. Besides water electrolysis, this electricity also supplies processes like \ce{CO2} hydrogenation \cite{brynolf2018,Decker2019,Pearson2012} or co-electrolysis of \ce{CO2} and \ce{H2O} \cite{hanggi2019,larsson2015}. Aside the two aforementioned keywords, the literature makes use of a scattered diversity of terms that aim at precising the features of the considered fuel: for instance, ``solar fuel", ``power gas" or ``green synthetic fuel". However, out of context, words like ``green fuel", ``advanced fuel" or ``solar fuel" may be misunderstood because of their multiple interpretations \cite{ridjan2016}.

@@

\textit{Electrofuels} are, according to \citet{JRC_alternative}, fuels whose energy content comes from renewable electricity. On top of that, \citet{goldmann2018} even add that they should be carbon-neutral and serve as a storage capacity of electricity. To be more comprehensive, we define \textit{electrofuels as fuels produced from electricity}, as the resource used to produce this electricity may vary. This electricity is transformed into hydrogen through electrolysis. This hydrogen can be upgraded in more complex fuels when combined with other molecules (\eg carbon mono/dioxide or nitrogen). Electrofuels group all the fuels where the absolute majority of the energy contained into the fuel comes from electricity. In this sense, it is important to point out that ``electrofuels" do not especially imply ``renewable fuels" (\Cref{fig:synfuels_paths}). For instance, the electricity needed to produce \ce{H2} through an electrolyser can be generated from renewable or non-renewable sources \cite{bhandari2014}, even if the most sustainable way to produce electrofuels remains through the electricity produced in excess from VRES.

>>>>

First and foremost, in the context of the energy transition, fuels will keep playing a major role in the global energy system \cite{Ahlgren2012}. Even if electricity gains shares through the electrification of the energy demand, it will not entirely displace fuels for three main reasons: storage, infrastructure compatibility and cross-sectorial links. Because of their intermittency and space disparity, a deeper integration of \gls{VRES} requires storage and transport in order to supply the energy demand at the right time and in the right place \cite{Brouwer2016,Evans2012,Gallo2016,hall2008,Rosa2017}. Where typical container of batteries are limited in terms of storage capacity (up to 10~MWh) and present significant costs and self-discharge losses, the energy conversion into fuels provides a more affordable solution for higher storage capacity (from 100~GWh) and longer storage time scales (months to year) \cite{Rosa2017}. Due to the economic inertia and their infrastructure legacy \cite{Ahlgren2012}, fuels remain the most appropriate solution for sectors requiring high energy density (\eg heavy-duty transport, shipping, aviation or chemical industry) \cite{Albrecht2020,Decker2019,Goede2018,Pearson2012,Rosa2017,Stancin2020,Trieb2018,Zeman2008}. 
\citet{contino2020whole} point out that the energy transition is an interdisciplinary effort and not solely about the power sector. The latter represents only a fifth of the global energy consumption \cite{iea_renewable}. Also, \citet{Goede2018} has shown in 2018 that \ce{CO2} emissions of the Netherlands are equally apportioned between the different types of end-use-demands (\ie power, heat, mobility and non-energy). This highlights the necessity to consider every energy sector rather than focusing all efforts on the power system and even more, to shift towards a multi-vector interconnected energy system. In this cross-sectorial approach and with the perspective of increasing shares of \gls{VRES}, fuels are promising energy carriers in order to maximise the overall system efficiency \cite{mathiesen2015, Stancin2020}.

@@

Where terms like ``synthetic gas", ``electrofuels" or ``synthetic natural gas" account for 19\%, 13\% and 8\% of the additional relevant literature, respectively, ``synthetic fuels" is used in 35\% of the case to cover fuels produced from a wide variety of primary resources (renewable or not) and through various conversion processes. To pick only a few examples, its usage ranged from liquid hydrocarbons produced from synthesis gas through the Fischer-Tropsch process \cite{haarlemmer2014,Trieb2018} to ammonia from the Haber-Bosch process \cite{bargiacchi2019} or biomass gasification or pyrolysis \cite{monaco2018, rao2015}. As a consequence of an increasing share of \gls{VRES} in the energy system, ``electrofuels" are gaining more popularity: where this term was used in 4\% of the relevant literature between 2006 and 2014 \cite{ridjan2016}, it is now present in 13\% of the additional relevant literature reviewed for this work. These fuels are considered as being produced mainly based on the electricity provided by \gls{VRES} \cite{brynolf2018}. Besides water electrolysis, this electricity also supplies processes like \ce{CO2} hydrogenation \cite{brynolf2018,Decker2019,Pearson2012} or co-electrolysis of \ce{CO2} and \ce{H2O} \cite{hanggi2019,larsson2015}. Aside the two aforementioned keywords, the literature makes use of a scattered diversity of terms that aim at precising the features of the considered fuel: for instance, ``solar fuel", ``power gas" or ``green synthetic fuel". However, out of context, words like ``green fuel", ``advanced fuel" or ``solar fuel" may be misunderstood because of their multiple interpretations \cite{ridjan2016}.

@@

\textit{Electrofuels} are, according to \citet{JRC_alternative}, fuels whose energy content comes from renewable electricity. On top of that, \citet{goldmann2018} even add that they should be carbon-neutral and serve as a storage capacity of electricity. To be more comprehensive, we define \textit{electrofuels as fuels produced from electricity}, as the resource used to produce this electricity may vary. This electricity is transformed into hydrogen through electrolysis. This hydrogen can be upgraded in more complex fuels when combined with other molecules (\eg carbon mono/dioxide or nitrogen). Electrofuels group all the fuels where the absolute majority of the energy contained into the fuel comes from electricity. In this sense, it is important to point out that ``electrofuels" do not especially imply ``renewable fuels" (\Cref{fig:synfuels_paths}). For instance, the electricity needed to produce \ce{H2} through an electrolyser can be generated from renewable or non-renewable sources \cite{bhandari2014}, even if the most sustainable way to produce electrofuels remains through the electricity produced in excess from \gls{VRES}.

### 2024-02-20 chap_intro_ccl/intro.tex
Merge taxonomy paper + Implement Hervé's comments on case_study
First part of intro: transition, VRES, electrofuels, ESOMs

@@

General introduction to the need of energy transition\\

How electrofuels will be part of the solution. Here include part of the terminology paper \url{https://www.frontiersin.org/articles/10.3389/fenrg.2021.660073/full} to clearly set the context and make sure the readers understand what's behind electrofuels and emphasize at the end that in our case, we consider electrofuels as the ones from renewable sources (with a 0-gwp).\\

>>>>

Consequently, to ensure the energy supply of an assumably more and more demanding society in a context of environmental crisis, major transformations are needed (see Figure \ref{fig:intro:IEA_WEO_2019}). Besides behavioral changes, an overall reshape of the energy system is necessary in terms of both primary energy sources, \ie more renewables, and technologies used to convert these resources into the \gls{EUD} (\ie the energy service required by the the final consumer), \ie more efficiency \cite{iea2020world}. The former corresponds to a whole ``fuel switch'' (see Figure \ref{fig:intro:IEA_WEO_2019}) where energy carriers called, in the literature, ``\emph{biofuels}'', ``\emph{electrofuels}'', ``\emph{synthetic fuels}'', ``\emph{renewable fuels}'' or even ``\emph{sustainable fuels}'', will more and more play a crucial role. To avoid the confusion between these fuels and thereby reduce misunderstanding in political or academic discussions, we have suggested a comprehensive and harmonised taxonomy (see Appendix \ref{app:Taxonomy}). In the rest of this thesis, the electro - and bio - fuels are considered as renewable and with no \gls{GWP}.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.9\textwidth]{IEA_WEO_2019.png}
\caption{Energy-related CO2 emissions and reductions by sources in the Sustainable Development Scenario \cite{iea2020world}.}
\label{fig:intro:IEA_WEO_2019}
\end{figure}

In the general perspective to decrease the \gls{GWP} of the primary energy mix, \gls{VRES} like wind and solar, have already emerged as the keystone to defossilise the energy system. However, their intermittency and space disparity could hold back their vaster integration in the future. To address this issue, due to some limitations (\eg range, power, costs) of electricity-focused solutions like \gls{DC} lines, the transport and long-term storage of the renewable electricity produced in excess should be optimised.

This challenge can be tackled by \textit{electrofuels} \cite{rozzi2020}. These fuels represent energy carriers where electricity has the major share in the energy balance of the fuel. In practice, this electricity is mainly converted into hydrogen (\ie electrolysis) and then potentially upgraded into more complex fuels (\eg methane, methanol or ammonia). Even if the share of electricity increases in the energy system through the electrification of the end-use demand, gaseous and liquid fuels will keep on being big players during (and after) the energy transition \cite{Ahlgren2012}. They offer three main advantages: infrastructure compatibility, storage and capacity to link sectors (\ie from electricity to mobility, heat, or industry). Development on electrofuels aims at getting them more and more compatible with existing and mature technologies \cite{Ahlgren2012}. An example is carbon-free ammonia-hydrogen blends burned in spark ignition engines \cite{lhuillier2020experimental} or \gls{CHP} applications \cite{pochet202022}. With a growing share of \gls{VRES}, sector coupling is essential to absorb the surplus of electricity from these intermittent production means \cite{robinius2017linking} and integrate them more cost-effectively \cite{brown2018response, limpensECOS2021}. Besides direct electrification of other sectors (\eg electrical heat pumps, battery electric vehicles), \citet{brown2018synergies} showed that converting power to hydrogen and methane was advantageous at high shares of renewables, in their optimisation of the European whole-energy system. Electrofuels have the ability to couple energy and non-energy sectors \cite{Stancin2020}. For instance, electricity produced in excess from \gls{VRES} can be converted in ammonia through the Haber-Bosch process and subsequently transformed into fertiliser - coupling the power and industry sectors \cite{verleysen2020can}. Gas networks present much more storage potential than electrical network (\eg 50 times more in Germany and 300 times more in France) \cite{Rosa2017}. Where batteries exhibit limited storage capacity (up to 10\,MWh) as well as self-discharge losses, electrofuels are an economical solution for high capacity (from 100 GWh) and long-term (\ie from months to years) storage of energy \cite{child2018role, dias2020energy} (see Figure \ref{fig:intro:Storage_electrofuels}.) Besides storing energy, in their analysis of the German transport sector in 2050, \citet{millinger2021electrofuels} highlighted that producing electrofuels can represent a better usage of the ambient \ce{CO2} than \gls{CCS} to supply hydrocarbon fuels while limiting the curtailment of \gls{VRES}. Moreover, some applications (\eg marine, aviation and heavy-duty transport) will be hard to electrify and keep on requiring high-density energy carriers \cite{horvath2018techno, brynolf2018}. These carriers, currently produced mostly from fossil resources, will still consist of hydrocarbons in a renewable world. This is why this paper rather uses "defossilisation" rather than "decarbonisation" as carbon will still play a key role in a carbon-neutral energy transition \cite{mertens2020carbon}.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.9\textwidth]{Storage_electrofuels.pdf}
\caption{Energy carriers and technologies to store electricity. Electrofuels are an economical solution for high capacity and long-term storage of energy. Graph adapted from \cite{ISPT2017}.}
\label{fig:intro:Storage_electrofuels}
\end{figure}

To harvest the maximum potential of synthetic energy carriers in a sustainable transition and maximise the overall system efficiency \cite{mathiesen2015}, it is necessary to study the integration of these fuels within a multi-sector and whole-energy system \cite{Contino2020}. To reach this goal, an energy system optimisation model (ESOM) can define the design of the system to minimise, for instance, its costs or its emissions \cite{zeng2011review}. In this research field, \citet{yue2018review} highlighted that most of ESOMs use a deterministic approach (\ie 75\% out of the 134 reviewed ESOM studies). However, the model structures are inherently uncertain as well as their numerous composing parameters, especially when it comes to define an energy transition strategy for a large-scale system, such as a country. Given the lifetime of the conversion technologies, such strategy implies decisions with long-term impacts (20 to 50 years) where forecasts can be highly unreliable \cite{Moret2017}. Besides the uncertainty on the model structure (not addressed in this work), this long-term and large-system optimisation motivates the need to account for \gls{UQ} and consider it as a major challenge of such models \cite{pfenninger2014energy}. This challenge, along with a large number (\ie more than a hundred) of uncertain parameters and limited information of their distribution, leads to the "curse of dimensionality" \cite{kuo2005lifting}.

@@
